---
title: "DataManagementInR"
author: "Kyle Black"
date: "2023-05-02"
output: html_document
---

```{r cars}
library("tidyverse")
mpg
```
#Loaded tidyverse package, viewed the MPG dataset.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x= displ, y= hwy, colour= class))
```
#Made our first plot, if you don't point a geom_point then you will end up with a blank plot. The colour aesthetic assigns each value of a variable a different color.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, size= class))
```
#The size aesthetic assigns each value of a variable a different size.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, alpha= class))
```
#The alpha aesthetic assigns each value of a variable a different transparency.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, shape= class))
```
#The shape aesthetic assigns each value of a variable a different shape.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy), color = "blue")
```
#Color = "x" makes all of your points one color.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, colour= displ < 5))
```
#The colour aesthetic in this line of code gives a certain colour to a value depending on if it meets the threshold (in this case, smaller than 5).

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy)) +
  facet_wrap(~class, nrow= 3)
```
#Facet wrap allows you to plot single variables using ~ you can select each variable is used to make the plot. ~Class plots each class individually, nrow determines the number of rows your plots are displayed in. Ncol would do the same thing but for columns.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x= displ, y= hwy)) +
  facet_grid(drv ~ cyl)
```
#Facet_grid can be used to plot a combination of two variables.

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x= displ, y = hwy))
```
#Geom_smooth displays the data as a smooth line through the data points.

```{r}
ggplot (data = mpg) +
  geom_smooth(mapping = aes(x= displ, y= hwy, colour = drv, linetype = drv))
```
#You can use linetype to separate values of a variable into different types of lines.

```{r}
ggplot (data= mpg) + 
  geom_smooth(mapping = aes(x= displ, y = hwy, color = drv))
```
#Displays the different values of the drive train variable as different colors. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```
#You can put your aesthetics with the ggplot argument so you don't have to repeat the same information in the geom_point and geom_smooth arguments.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes( colour = class)) +
  geom_smooth()
```
#You can also add different aesthetics to different layers of a plot while still using the ggplot argument to set up your plot.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```
#You can also specify different data for each layer.

```{r}
ggplot (data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
#Visualize the diamonds dataset.

```{r}
ggplot (data = diamonds) +
  stat_count(mapping = aes(x = cut))
```
#You can use geoms and stats interchangeably.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```
#Displays a bar chart of a proportion (hence prop) of your total diamond dataset instead of a count. Group = 1 does xyz!!!!!!!!!!!!!!!!!

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
#Plots statistical transformations

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```
#Colour and fill can both be used to better display information from aesthetics.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
#The stacking of fill colors is automatically done by ggplot.

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 2/5, position = "identity")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 3/5, position = "fill")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 2/5, position = "dodge")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 4/5, position = "jitter")
```
#Position adjustments allow us to customize our plots; identity (raw data), fill (changes heights), dodge (forces ggplot to not put things on top of each other), and jitter (shifts points a small random amount to avoid overlapping, not useful for bar charts. 
```{r}
ggplot(data = <DATA>) +
  <geom_function>
  (mapping = aes(x=, y=)),
  stat = <STAT>,
  position = <POSITION> +
  FACET_FUNCTION
```
#GGplot template for future graphs

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(title = "Fuel efficieny generally decreases with engine size.")
```
#You can use labs() to create labels for your plot. Title, subtitle, and caption can all be used to add descriptive texts to your plot.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(title = "Fuel efficieny generally decreases with engine size.", subtitle = "Two seaters (sports cars) are an exception because of their light weight.", 
  caption = "Data from fueleconomy.gov")
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(x = "Engine displacement (L)", y = "Highway fuel economy (mpg)", Colour = "Car type")
```
#You can also use labs() to name axsis titles.

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_text(aes(label = model), data = best_in_class)
```
#Annotations are useful to highlight specific observations in your data. Nudge() can be used to move overlapping labels.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(limits = c(0, 8)) +
  scale_y_continuous(limits = c(0, 45)) +
  scale_color_discrete()
```
#Scale continuous allows you to modify the scale of the axes of your plot. Scale_color_discrete() does skjdbfskjfdkjgkgjbejeff

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15,40, by = 5))
```
#You can change the tick marks on your plot's axes.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```
#You can turn off/avoid labels altogether by making them NULL.

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right")
base + theme(legend.position = "none")
```
#You can change the position of your figure's legend to make it look more aesthetically pleasing. Or remove it completely!

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() +
  scale_x_log10() +
  scale_y_log10()
```
#Log transforms the x, y scales

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set3")
```
#Changes the color/pattern scheme for the plot.

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```
#

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x,y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```
#Trying out the viridis colour scheme.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_light()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_dark()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_classic()
```
#GGplot has different default themes you can use to change some or all of the elements of your figure.

```{r}
theme (panel.border = element_blank(),
       panel.grid.minor.x = element_blank(),
       panel.grid.minor.y = element_blank(),
       legend.position = "right",
       legend.title = element_blank(),
       legend.text = element_text(size = 8),
       panel.grid.major = element_blank(),
       legend.key = element_blank(),
       legend.background = element_blank(),
       axis.text.y = element_text(colour="black"),
       axis.text.x = elemenet_text(colour= "black"),
       text = element_text(family = "Arial")
       )
```
#You can also develop your own theme! 


#Workshop 4 - Data wrangling in R
```{r}
as_tibble(iris)
```
#You can convert a dataset to a tibble using the function as_tibble()

```{r}
tibble(
  x = 1:5,
  y = 1,
  z = x^2 + y
)
```
#You can call new variables to produce new column values.

```{r}
tb <- tibble(
  `:)` = "smile",
  `:(` = "frown",
  `<3` = "kaysha"
)

tb
```


```{r}
tribble(
  ~x, ~y, ~z,
  #---|---|---
  "a", 2, 3.6,
  "b", 1, 8.5
)
```


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```
#Date formats in a tibble

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```


```{r}
options(tibble.print_min = Inf) #Shows all rows
options(tibble.width = Inf) #Print all rows
```


```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x
df[["x"]]
#Extracting data by name

df[[1]]
#Extracting data by row position

df[[3,2]]
#Extract by exact position [row, column]
```


```{r}
df <- data.frame(abc = 1, xyz = "a")
df
df$x #call by name!!!!!!!!!!!!!!!!!!!!!!!!!!
df[, "xyz"] #call by exact position
```


```{r}
read_csv("a,b,c
         1,2,3
         4,5,6"
         )

read_csv("The first line of metadata
         The second line of metadata
         x, y, z
         1, 2, 3", skip = 2)

read_csv("#A comment I want to skip
         x, y, z
         1, 2, 3", comment = '#')

read_csv("1, 2, 3\n4, 5, 6", col_names= FALSE)

read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```
#You can use read functions to read in your data if you know the file path. Otherwise you can create your data in R by typing it in manually. 
#You can use skip = n to skip a certain number of lines. #Additionally, you can use comment = '#' to remove all lines that begin with a # (or you can use another character if you so choose).
#You can use col_names = FALSE to not treat the first row of data as headings.\n is a shortcut for adding a new line
#Col_names can be used to name columns as you see fit.

```{r}
table1 #An example of tidy data
```
#Rules of tidy data:
#1) Each variable must have its own column.
#2) Each observation must have its own row.
#3 Each value must have its own cell.

```{r}
table1 %>% mutate(rate = cases / population * 10000) #Compute rate per 10,000 people

table1 %>% count(year, wt = cases) #Compute cases per year

ggplot(table1, aes(year, cases)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country))
```


```{r}
pivot_longer() 5.5.2

pivot_wider()
```