---
title: "DataManagementInR"
author: "Kyle Black"
date: "2023-05-02"
output: html_document
---

```{r cars}
library(tidyverse)
library(ggplot2)
library(sf)
library(terra)
library(tmap)
```
#Load packages

```{r}
mpg
```
#View the MPG dataset

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x= displ, y= hwy, colour= class))
```
#Made our first plot, if you don't point a geom_point then you will end up with a blank plot. The colour aesthetic assigns each value of a variable a different color.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, size= class))
```
#The size aesthetic assigns each value of a variable a different size.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, alpha= class))
```
#The alpha aesthetic assigns each value of a variable a different transparency.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, shape= class))
```
#The shape aesthetic assigns each value of a variable a different shape.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy), color = "blue")
```
#Color = "x" makes all of your points one color.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, colour= displ < 5))
```
#The colour aesthetic in this line of code gives a certain colour to a value depending on if it meets the threshold (in this case, smaller than 5).

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy)) +
  facet_wrap(~class, nrow= 3)
```
#Facet wrap allows you to plot single variables using ~ you can select each variable is used to make the plot. ~Class plots each class individually, nrow determines the number of rows your plots are displayed in. Ncol would do the same thing but for columns.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x= displ, y= hwy)) +
  facet_grid(drv ~ cyl)
```
#Facet_grid can be used to plot a combination of two variables.

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x= displ, y = hwy))
```
#Geom_smooth displays the data as a smooth line through the data points.

```{r}
ggplot (data = mpg) +
  geom_smooth(mapping = aes(x= displ, y= hwy, colour = drv, linetype = drv))
```
#You can use linetype to separate values of a variable into different types of lines.

```{r}
ggplot (data= mpg) + 
  geom_smooth(mapping = aes(x= displ, y = hwy, color = drv))
```
#Displays the different values of the drive train variable as different colors. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```
#You can put your aesthetics with the ggplot argument so you don't have to repeat the same information in the geom_point and geom_smooth arguments.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes( colour = class)) +
  geom_smooth()
```
#You can also add different aesthetics to different layers of a plot while still using the ggplot argument to set up your plot.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```
#You can also specify different data for each layer.

```{r}
ggplot (data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
#Visualize the diamonds dataset.

```{r}
ggplot (data = diamonds) +
  stat_count(mapping = aes(x = cut))
```
#You can use geoms and stats interchangeably.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```
#Displays a bar chart of a proportion (hence prop) of your total diamond dataset instead of a count. Group = 1 does xyz!!!!!!!!!!!!!!!!!

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
#Plots statistical transformations

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```
#Colour and fill can both be used to better display information from aesthetics.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
#The stacking of fill colors is automatically done by ggplot.

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 2/5, position = "identity")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 3/5, position = "fill")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 2/5, position = "dodge")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 4/5, position = "jitter")
```
#Position adjustments allow us to customize our plots; identity (raw data), fill (changes heights), dodge (forces ggplot to not put things on top of each other), and jitter (shifts points a small random amount to avoid overlapping, not useful for bar charts. 
```{r}
ggplot(data = <DATA>) +
  <geom_function>
  (mapping = aes(x=, y=)),
  stat = <STAT>,
  position = <POSITION> +
  FACET_FUNCTION
```
#GGplot template for future graphs

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(title = "Fuel efficieny generally decreases with engine size.")
```
#You can use labs() to create labels for your plot. Title, subtitle, and caption can all be used to add descriptive texts to your plot.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(title = "Fuel efficieny generally decreases with engine size.", subtitle = "Two seaters (sports cars) are an exception because of their light weight.", 
  caption = "Data from fueleconomy.gov")
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(x = "Engine displacement (L)", y = "Highway fuel economy (mpg)", Colour = "Car type")
```
#You can also use labs() to name axsis titles.

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_text(aes(label = model), data = best_in_class)
```
#Annotations are useful to highlight specific observations in your data. Nudge() can be used to move overlapping labels.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(limits = c(0, 8)) +
  scale_y_continuous(limits = c(0, 45)) +
  scale_color_discrete()
```
#Scale continuous allows you to modify the scale of the axes of your plot. Scale_color_discrete() does skjdbfskjfdkjgkgjbejeff

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15,40, by = 5))
```
#You can change the tick marks on your plot's axes.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```
#You can turn off/avoid labels altogether by making them NULL.

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right")
base + theme(legend.position = "none")
```
#You can change the position of your figure's legend to make it look more aesthetically pleasing. Or remove it completely!

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() +
  scale_x_log10() +
  scale_y_log10()
```
#Log transforms the x, y scales

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set3")
```
#Changes the color/pattern scheme for the plot.

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```
#

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x,y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```
#Trying out the viridis colour scheme.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_light()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_dark()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_classic()
```
#GGplot has different default themes you can use to change some or all of the elements of your figure.

```{r}
theme (panel.border = element_blank(),
       panel.grid.minor.x = element_blank(),
       panel.grid.minor.y = element_blank(),
       legend.position = "right",
       legend.title = element_blank(),
       legend.text = element_text(size = 8),
       panel.grid.major = element_blank(),
       legend.key = element_blank(),
       legend.background = element_blank(),
       axis.text.y = element_text(colour="black"),
       axis.text.x = elemenet_text(colour= "black"),
       text = element_text(family = "Arial")
       )
```
#You can also develop your own theme! 


#Workshop 4 - Data wrangling in R
```{r}
as_tibble(iris)
```
#You can convert a dataset to a tibble using the function as_tibble()

```{r}
tibble(
  x = 1:5,
  y = 1,
  z = x^2 + y
)
```
#You can call new variables to produce new column values.

```{r}
tb <- tibble(
  `:)` = "smile",
  `:(` = "frown",
  `<3` = "kaysha"
)

tb
```


```{r}
tribble(
  ~x, ~y, ~z,
  #---|---|---
  "a", 2, 3.6,
  "b", 1, 8.5
)
```


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```
#Date formats in a tibble

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```


```{r}
options(tibble.print_min = Inf) #Shows all rows
options(tibble.width = Inf) #Print all rows
```


```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x
df[["x"]]
#Extracting data by name

df[[1]]
#Extracting data by row position

df[[3,2]]
#Extract by exact position [row, column]
```


```{r}
df <- data.frame(abc = 1, xyz = "a")
df
df$x #call by name!!!!!!!!!!!!!!!!!!!!!!!!!!
df[, "xyz"] #call by exact position
```


```{r}
read_csv("a,b,c
         1,2,3
         4,5,6"
         )

read_csv("The first line of metadata
         The second line of metadata
         x, y, z
         1, 2, 3", skip = 2)

read_csv("#A comment I want to skip
         x, y, z
         1, 2, 3", comment = '#')

read_csv("1, 2, 3\n4, 5, 6", col_names= FALSE)

read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```
#You can use read functions to read in your data if you know the file path. Otherwise you can create your data in R by typing it in manually. 
#You can use skip = n to skip a certain number of lines. #Additionally, you can use comment = '#' to remove all lines that begin with a # (or you can use another character if you so choose).
#You can use col_names = FALSE to not treat the first row of data as headings.\n is a shortcut for adding a new line
#Col_names can be used to name columns as you see fit.

```{r}
table1 #An example of tidy data
```
#Rules of tidy data:
#1) Each variable must have its own column.
#2) Each observation must have its own row.
#3 Each value must have its own cell.

```{r}
table1 %>% mutate(rate = cases / population * 10000) #Compute rate per 10,000 people

table1 %>% count(year, wt = cases) #Compute cases per year

ggplot(table1, aes(year, cases)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country))
```


```{r}
table4a %>% pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")

table4b %>% pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
```
#Pivot longer increases the number of rows and decreases the number of columns. It is also used to rename columns that have values in place of column names.

```{r}
table2 %>%
    pivot_wider(names_from = type, values_from = count)
```
#Pivot wider makes the table shorter and wider by creating new columns.

```{r}
table3

table3 %>% 
  separate(rate, into = c("cases", "population"))
```
#Splits the "rate" column into 2 different columns and renames the old and names the new column.

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
```
#Converted the columns cases and population from characters into a more appropriate class.

```{r}
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```
#Separated the column year into the century and year.


```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```
#Unite is the opposite for separate, and combines the century and year columns.

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)

stocks %>% 
  pivot_wider(names_from = year, values_from = return)

stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(
    cols = c(`2015`, `2016`), 
    names_to = "year", 
    values_to = "return", 
    values_drop_na = TRUE
  )

```
#Removes values without any data from the table by creating the year columns and dropping any NA values.

```{r}
stocks %>% 
  complete(year, qtr)
```
#Complete takes a set of columns and finds all of the unique combinatinos and makes sure the original datasets contains all of the values. This includes any NA values where they are needed.

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)
treatment

treatment %>% 
  fill(person)
```
#Fill is used to fill in any missing data entries. Takes columns with missig values and carries the last non-missing value forward into the areas where values are missing.

```{r}
#5.6 here, optional but do it!!!!!!!
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}
#install.packages("sf")
#install.packages("terra")
#install.packages("tmap")
```
#These packages are for using spatial data

```{r}
dat <- read_csv("R/Data/data-for-course/copepods_raw.csv")
dat
```

```{r}
ggplot(dat) +
  aes(x = longitude, y = latitude, color = richness_raw) +
  geom_point()
#This first "map" lacks all the features of a map including projection (bend of your data over the spherical globe)

ggplot(dat, aes(x = latitude, y = richness_raw)) + 
  stat_smooth() + 
  geom_point()
```

```{r}
sdat <- st_as_sf(dat, coords = c("longitude", "latitude"), 
                 crs = 4326)
```
#Turn point data into spatially referenced data using R's packages for GIS and mapping.

```{r}
crs4326 <- st_crs(4326)
crs4326 
crs4326$Name #Pulls out the whole coordinate reference system (CRS)
crs4326$wkt #CRS in well-known ext (WKT) format
```

```{r}
sdat
```

```{r}
plot(sdat["richness_raw"])
```
#You can use square brackets to plot a single variable.

```{r}
tm_shape(sdat) + 
  tm_dots(col = "richness_raw")
```
#tm_dots plots coordinates as dot points, other options include tm_polygons and tm_symbols

```{r}
tmap_save(filename = "Richness-map.png", 
          width = 600, height = 600)
```
#Tmap_save saves a copy of your map to your current working directory.

```{r}
aus <- st_read("R/Data/data-for-course/spatial-data/Aussie/Aussie.shp")
shelf <- st_read("R/Data/data-for-course/spatial-data/aus_shelf/aus_shelf.shp")
```
#Reads shapefiles into R

```{r}
tm_shape(shelf) + 
  tm_polygons()
```
#Tmap provides an easy way to make a map of polygons

```{r}
tm_shape(shelf, bbox = sdat) + 
  tm_polygons() +
  tm_shape(aus) + 
  tm_polygons() + 
  tm_shape(sdat) + 
  tm_dots() +
  tmap_style("albatross")
```
#Like ggplot, tmap allows you to add different layers onto your map.

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```



