---
title: "Module 4 Workshop 1"
author: "Kyle Black"
date: "2023-05-19"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(ggplot2)
library(sf)
library(terra)
library(tmap)
```

#Load packages

```{r}
mpg
```

#View the MPG dataset

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x= displ, y= hwy, colour= class))
```

#Made our first plot, if you don't point a geom_point then you will end up with a blank plot. The colour aesthetic assigns each value of a variable a different color.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, size= class))
```

#The size aesthetic assigns each value of a variable a different size.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, alpha= class))
```

#The alpha aesthetic assigns each value of a variable a different transparency.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, shape= class))
```

#The shape aesthetic assigns each value of a variable a different shape.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy), color = "blue")
```

#Color = "x" makes all of your points one color.

```{r}
ggplot(data= mpg) + 
  geom_point(mapping = aes(x= displ, y = hwy, colour= displ < 5))
```

#The colour aesthetic in this line of code gives a certain colour to a value depending on if it meets the threshold (in this case, smaller than 5).

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy)) +
  facet_wrap(~class, nrow= 3)
```

#Facet wrap allows you to plot single variables using ~ you can select each variable is used to make the plot. ~Class plots each class individually, nrow determines the number of rows your plots are displayed in. Ncol would do the same thing but for columns.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x= displ, y= hwy)) +
  facet_grid(drv ~ cyl)
```

#Facet_grid can be used to plot a combination of two variables.

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x= displ, y = hwy))
```

#Geom_smooth displays the data as a smooth line through the data points.

```{r}
ggplot (data = mpg) +
  geom_smooth(mapping = aes(x= displ, y= hwy, colour = drv, linetype = drv))
```

#You can use linetype to separate values of a variable into different types of lines.

```{r}
ggplot (data= mpg) + 
  geom_smooth(mapping = aes(x= displ, y = hwy, color = drv))
```

#Displays the different values of the drive train variable as different colors. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```

#You can put your aesthetics with the ggplot argument so you don't have to repeat the same information in the geom_point and geom_smooth arguments.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes( colour = class)) +
  geom_smooth()
```

#You can also add different aesthetics to different layers of a plot while still using the ggplot argument to set up your plot.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

#You can also specify different data for each layer.

```{r}
ggplot (data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

#Visualize the diamonds dataset.

```{r}
ggplot (data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

#You can use geoms and stats interchangeably.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```

#Displays a bar chart of a proportion (hence prop) of your total diamond dataset instead of a count. Group = 1 does xyz!!!!!!!!!!!!!!!!!

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

#Plots statistical transformations

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

#Colour and fill can both be used to better display information from aesthetics.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

#The stacking of fill colors is automatically done by ggplot.

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 2/5, position = "identity")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 3/5, position = "fill")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 2/5, position = "dodge")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 4/5, position = "jitter")
```

#Position adjustments allow us to customize our plots; identity (raw data), fill (changes heights), dodge (forces ggplot to not put things on top of each other), and jitter (shifts points a small random amount to avoid overlapping, not useful for bar charts. 
```{r, eval=FALSE}
ggplot(data = <DATA>) +
  <geom_function>
  (mapping = aes(x=, y=)),
  stat = <STAT>,
  position = <POSITION> +
  FACET_FUNCTION
```

#GGplot template for future graphs