---
title: "Module 4 Workshop 4"
author: "Kyle Black"
date: "2023-05-19"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(ggplot2)
library(sf)
library(terra)
library(tmap)
```

#Load packages

```{r}
as_tibble(iris)
```

#You can convert a dataset to a tibble using the function as_tibble()

```{r}
tibble(
  x = 1:5,
  y = 1,
  z = x^2 + y
)
```

#You can call new variables to produce new column values.

```{r}
tb <- tibble(
  `:)` = "smile",
  `:(` = "frown",
  `<3` = "kaysha"
)

tb
```


```{r}
tribble(
  ~x, ~y, ~z,
  #---|---|---
  "a", 2, 3.6,
  "b", 1, 8.5
)
```


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

#Date formats in a tibble

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```


```{r}
options(tibble.print_min = Inf) #Shows all rows
options(tibble.width = Inf) #Print all rows
```


```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x
df[["x"]]
#Extracting data by name

df[[1]]
#Extracting data by row position

df[[3,2]]
#Extract by exact position [row, column]
```


```{r}
df <- data.frame(abc = 1, xyz = "a")
df
df$x #call by name!!!!!!!!!!!!!!!!!!!!!!!!!!
df[, "xyz"] #call by exact position
```


```{r}
read_csv("a,b,c
         1,2,3
         4,5,6"
         )

read_csv("The first line of metadata
         The second line of metadata
         x, y, z
         1, 2, 3", skip = 2)

read_csv("#A comment I want to skip
         x, y, z
         1, 2, 3", comment = '#')

read_csv("1, 2, 3\n4, 5, 6", col_names= FALSE)

read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

#You can use read functions to read in your data if you know the file path. Otherwise you can create your data in R by typing it in manually. 

#You can use skip = n to skip a certain number of lines. #Additionally, you can use comment = '#' to remove all lines that begin with a # (or you can use another character if you so choose).

#You can use col_names = FALSE to not treat the first row of data as headings.\n is a shortcut for adding a new line

#Col_names can be used to name columns as you see fit.

```{r}
table1 #An example of tidy data
```
#Rules of tidy data:
#1) Each variable must have its own column.
#2) Each observation must have its own row.
#3 Each value must have its own cell.

```{r}
table1 %>% mutate(rate = cases / population * 10000) #Compute rate per 10,000 people

table1 %>% count(year, wt = cases) #Compute cases per year

ggplot(table1, aes(year, cases)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country))
```


```{r}
table4a %>% pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")

table4b %>% pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
```

#Pivot longer increases the number of rows and decreases the number of columns. It is also used to rename columns that have values in place of column names.

```{r}
table2 %>%
    pivot_wider(names_from = type, values_from = count)
```

#Pivot wider makes the table shorter and wider by creating new columns.

```{r}
table3

table3 %>% 
  separate(rate, into = c("cases", "population"))
```

#Splits the "rate" column into 2 different columns and renames the old and names the new column.

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
```

#Converted the columns cases and population from characters into a more appropriate class.

```{r}
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```

#Separated the column year into the century and year.


```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```

#Unite is the opposite for separate, and combines the century and year columns.

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)

stocks %>% 
  pivot_wider(names_from = year, values_from = return)

stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(
    cols = c(`2015`, `2016`), 
    names_to = "year", 
    values_to = "return", 
    values_drop_na = TRUE
  )
```

#Removes values without any data from the table by creating the year columns and dropping any NA values.

```{r}
stocks %>% 
  complete(year, qtr)
```

#Complete takes a set of columns and finds all of the unique combinatinos and makes sure the original datasets contains all of the values. This includes any NA values where they are needed.

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)
treatment

treatment %>% 
  fill(person)
```

#Fill is used to fill in any missing data entries. Takes columns with missig values and carries the last non-missing value forward into the areas where values are missing.