---
title: "Module 4 Workshop 2"
author: "Kyle Black"
date: "2023-05-19"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(ggplot2)
library(sf)
library(terra)
library(tmap)
```

#Load packages

```{r}
mpg
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(title = "Fuel efficieny generally decreases with engine size.")
```

#You can use labs() to create labels for your plot. Title, subtitle, and caption can all be used to add descriptive texts to your plot.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(title = "Fuel efficieny generally decreases with engine size.", subtitle = "Two seaters (sports cars) are an exception because of their light weight.", 
  caption = "Data from fueleconomy.gov")
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) + 
  labs(x = "Engine displacement (L)", y = "Highway fuel economy (mpg)", Colour = "Car type")
```

#You can also use labs() to name axsis titles.

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_text(aes(label = model), data = best_in_class)
```

#Annotations are useful to highlight specific observations in your data. Nudge() can be used to move overlapping labels.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(limits = c(0, 8)) +
  scale_y_continuous(limits = c(0, 45)) +
  scale_color_discrete()
```

#Scale continuous allows you to modify the scale of the axes of your plot. Scale_color_discrete() does skjdbfskjfdkjgkgjbejeff

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15,40, by = 5))
```

#You can change the tick marks on your plot's axes.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```

#You can turn off/avoid labels altogether by making them NULL.

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right")
base + theme(legend.position = "none")
```

#You can change the position of your figure's legend to make it look more aesthetically pleasing. Or remove it completely!

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() +
  scale_x_log10() +
  scale_y_log10()
```

#Log transforms the x, y scales

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set3")
```

#Changes the color/pattern scheme for the plot.

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```


```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x,y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```

#Trying out the viridis colour scheme.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_light()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_dark()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se= FALSE) +
  theme_classic()
```

#GGplot has different default themes you can use to change some or all of the elements of your figure.

```{r, eval=FALSE}
theme (panel.border = element_blank(),
       panel.grid.minor.x = element_blank(),
       panel.grid.minor.y = element_blank(),
       legend.position = "right",
       legend.title = element_blank(),
       legend.text = element_text(size = 8),
       panel.grid.major = element_blank(),
       legend.key = element_blank(),
       legend.background = element_blank(),
       axis.text.y = element_text(colour="black"),
       axis.text.x = elemenet_text(colour= "black"),
       text = element_text(family = "Arial")
       )
```

#You can also develop your own theme! 